#Setting global options 
``` {r set_options}
    
    knitr::opts_chunk$set(warning = FALSE)
```

#Loading and preprocessing
```{r}
    


    activity<-read.csv("C:\\Users\\pmunj\\Desktop\\R codes\\activity.csv")

    activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
    weekday <- weekdays(activity$date)
    activity <- cbind(activity,weekday)

    summary(activity)
```    

# 1. Mean number of total steps taken per day
```{r}
    total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(total_steps) <- c("date", "steps")
hist(total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "red", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

mean total number of steps taken per day:
```{r}
    mean(total_steps$steps)
```
median of the total number of steps taken
```{r}
median(total_steps$steps)
```
#2. Average daily activity pattern 
```{r}
    avg_daily_activity <- aggregate(activity$steps,     by=list(activity$interval), FUN=mean, na.rm=TRUE)
    names(avg_daily_activity) <- c("interval", "mean")
    plot(avg_daily_activity$interval, avg_daily_activity$mean, type = "l", col="darkblue", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
    avg_daily_activity[which.max(avg_daily_activity$mean),]$interval
```
#3. Filling missing values
```{r}
    sum(is.na(activity$steps))
```
```{r}
   missing <- avg_daily_activity$mean[match(activity$interval, avg_daily_activity$interval)]
```

```{r}
    activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = missing, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```

plotting the histogram

```{r}
    hist(total_steps_imputed$daily_steps, col = "IndianRed", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```

mean of total number of steps taken per day
```{r}
    mean(total_steps_imputed$daily_steps)
```
median of total steps taken per day
```{r}
    median(total_steps_imputed$daily_steps)
```
#4. finding differences in acticity patterns between weekdays and weekends 

```{r}
    activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```
time series plot of the 5 minute interval and average number of steps taken
```{r}
    library(ggplot2)
    activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```






